---
title: "ESM 262 HW 2"
author: "Lauren Kaapcke"
date: "May 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import and Tidy

1. Load tidyverse and read gazetteer data into a tibble. 

```{r tools_data, warning = F, message = F}

library(tidyverse)
library(knitr)
library(DBI)
library(dplyr)

#Read in data
gaz_raw <- as_tibble(
  read_delim(
    "CA_Features_20180401.txt", 
    delim = "|", col_names = T, 
    col_types = cols(.default = col_character())))

```

2. Select desired columns.

```{r table, warning = F, message = F}

gaz <- gaz_raw %>% 
  select(FEATURE_ID:STATE_ALPHA, 
         COUNTY_NAME, 
         ends_with("_DEC"), 
         ELEV_IN_M, 
         MAP_NAME:DATE_EDITED)
gaz


```

3. Convert the columns to the appropriate type. Convert unknown placeholders to NA. 

```{r convert, warning = F, message = F}

# Convert columns
gaz$FEATURE_ID      = parse_integer(gaz$FEATURE_ID)
gaz$PRIM_LAT_DEC    = parse_double(gaz$PRIM_LAT_DEC)
gaz$PRIM_LONG_DEC   = parse_double(gaz$PRIM_LONG_DEC)
gaz$SOURCE_LAT_DEC  = parse_double(gaz$SOURCE_LAT_DEC)
gaz$SOURCE_LONG_DEC = parse_double(gaz$SOURCE_LONG_DEC)
gaz$ELEV_IN_M       = parse_integer(gaz$ELEV_IN_M)
gaz$DATE_CREATED    = parse_date(gaz$DATE_CREATED, format = "%m/%d/%Y")
gaz$DATE_EDITED     = parse_date(gaz$DATE_EDITED, format = "%m/%d/%Y")

# Specify NA values
is.na(gaz$MAP_NAME)      <- gaz$MAP_NAME == "Unknown"
is.na(gaz$PRIM_LAT_DEC)  <- gaz$PRIM_LAT_DEC == 0
is.na(gaz$PRIM_LONG_DEC) <- gaz$PRIM_LONG_DEC == 0

head(gaz) #Check that type conversions worked

```

4. Delete rows with NA for primary longitude or latitude or where the feature is not California.

```{r filter, message = F, warning = F}

# Filter for CA and remove NA values
gaz <- filter(gaz,
              STATE_ALPHA == "CA" &
              !is.na(PRIM_LAT_DEC) &
              !is.na(PRIM_LONG_DEC))
           
gaz

```


5. Create a connection to the database and add gaz.

```{r database, message = F, warning = F}

gaz.db <- DBI::dbConnect(RSQLite::SQLite(), path = ":memory:")

copy_to(gaz.db, gaz, "gaz", # What does the "::" mean?
  temporary = FALSE, 
  overwrite = TRUE,
  indexes = list(
    "FEATURE_ID", "FEATURE_NAME" 
  )
)

# Create a table cookie that connects to the database
gaz_db <- tbl(gaz.db, "gaz")
gaz_db

```

## Analyze the Data

1. What is the most frequenty appearing feature name?

```{r feature_name, message = F, warning = F}

max_name <- dbGetQuery(gaz.db, "SELECT FEATURE_NAME AS 'Feature Name', MAX(y.num) AS Count
  FROM (SELECT FEATURE_NAME, COUNT() AS num
    FROM gaz
    GROUP BY FEATURE_NAME) y;")

max_name

```

The feature name that appears the most often is "Church of Christ".

2. What is the least frequently occuring feature class?

```{r feature_class, message = F, warning = F}

count_class <- dbGetQuery(gaz.db, "SELECT FEATURE_CLASS AS 'Feature Class', COUNT() AS n_classes                             FROM gaz
                          GROUP BY FEATURE_CLASS
                          ORDER BY n_classes;")
count_class

# When filtering for min, it would only return 1 value :(

```

The feature classes that appear the least are "Sea" and "Isthmus".

3. What is the approximate center point of each county?

```{r center, warning = F, message = F}

center <- dbGetQuery(gaz.db, "SELECT 
                     COUNTY_NAME AS 'County Name',
                     (MIN(PRIM_LAT_DEC)+MAX(PRIM_LAT_DEC))/2 AS 'Center Latitude',
                     (MIN(PRIM_LONG_DEC)+MAX(PRIM_LONG_DEC))/2 AS 'Center Longitude'
                     FROM gaz 
                     WHERE COUNTY_NAME IS NOT NULL
                     GROUP BY COUNTY_NAME")
center
```

4. What are the fractions of the total number of features in each county that are natural vs. man-made?

```{r feature_type, message = F, warning = F}

# Isolate the unique feature classes
classes <- dbGetQuery(gaz.db,"SELECT FEATURE_CLASS FROM gaz
                      GROUP BY FEATURE_CLASS
                      ORDER BY FEATURE_CLASS;")

classes

# Add classes to the database
copy_to(gaz.db, classes, "classes_db",
  temporary = FALSE, 
  overwrite = TRUE
)

# Create a cookie to talk to the table
classes_db <- tbl(gaz.db, "classes_db")

# Create a new tibble that defines each class as natural or manmade
types <- mutate(classes,
                Class_Type = c("Manmade", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Manmade", "Manmade", "Manmade", "Natural", "Manmade", "Manmade", "Natural", "Manmade", "Manmade", "Natural", "Natural", "Manmade", "Manmade", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Manmade", "Manmade", "Natural", "Natural", "Natural", "Natural", "Natural", "Manmade", "Manmade", "Manmade", "Manmade", "Manmade", "Natural", "Natural", "Manmade", "Manmade", "Natural", "Natural", "Manmade", "Manmade", "Natural", "Manmade", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Manmade", "Manmade", "Manmade", "Natural", "Manmade", "Natural")
)

# Add types to the database
copy_to(gaz.db, types, "types_db",
  temporary = FALSE, 
  overwrite = TRUE
)

# Create a cookie to talk to the table
types_db <- tbl(gaz.db, "types_db")

# Join to gaz, select desired columns
class_types <- dbGetQuery(gaz.db, "SELECT gaz.FEATURE_NAME, gaz.FEATURE_CLASS, gaz.COUNTY_NAME, types_db.Class_Type FROM gaz JOIN types_db ON gaz.FEATURE_CLASS = types_db.FEATURE_CLASS;")

class_types

copy_to(gaz.db, class_types, "class_types_db",
  temporary = FALSE, 
  overwrite = TRUE
)

# Create a cookie to talk to the table
class_types_db <- tbl(gaz.db, "class_types_db")
class_types_db

# Pull the different metrics you need
manmade <- dbGetQuery(gaz.db, "SELECT COUNTY_NAME AS 'County', count(Class_Type) AS 'Manmade' FROM class_types_db WHERE Class_Type = 'Manmade' AND COUNTY_NAME != 'NA' GROUP BY COUNTY_NAME;")
manmade

# Add to the database
copy_to(gaz.db, manmade, "manmade_db",
  temporary = FALSE, 
  overwrite = TRUE
)

# Create a cookie to talk to the table
manmade_db <- tbl(gaz.db, "manmade_db")

natural <- dbGetQuery(gaz.db, "SELECT COUNTY_NAME AS 'County', count(Class_Type) AS 'Natural' FROM class_types_db WHERE Class_Type = 'Natural' AND COUNTY_NAME != 'NA' GROUP BY COUNTY_NAME;")

# Add to the database
copy_to(gaz.db, natural, "natural_db",
  temporary = FALSE, 
  overwrite = TRUE
)

# Create a cookie to talk to the table
natural_db <- tbl(gaz.db, "natural_db")

# Join these and calculate the fractions
fractions <- dbGetQuery(gaz.db, "SELECT manmade_db.County, CAST(manmade_db.Manmade AS REAL)/(manmade_db.Manmade + natural_db.Natural) AS 'Fraction Manmade', CAST(natural_db.Natural AS REAL)/(manmade_db.Manmade + natural_db.Natural) AS 'Fraction Natural' FROM manmade_db JOIN natural_db ON manmade_db.County = natural_db.County;")

fractions

```



