---
title: "ESM 262 HW 2"
author: "Lauren Kaapcke"
date: "May 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import and Tidy

1. Load tidyverse and read gazetteer data into a tibble. 

```{r tools_data, warning = F, message = F}

library(tidyverse)
library(knitr)
library(DBI)
library(dplyr)

#Read in data
gaz_raw <- as_tibble(read_delim("C:/Users/lnkaa/Documents/ESM Computing/CA_Features_20180401.txt", delim = "|", col_names = T, col_types = cols(.default = col_character())))

```

2. Select desired columns.

```{r table, warning = F, message = F}

gaz <- gaz_raw %>% 
  select(FEATURE_ID:STATE_ALPHA, COUNTY_NAME, ends_with("_DEC"), ELEV_IN_M, MAP_NAME:DATE_EDITED)

```

3. Convert the columns to the appropriate type. Convert unknown placeholders to NA. 

```{r convert, warning = F, message = F}

#Replace "Unknown" with "NA"
is.na(gaz) <- gaz == "Unknown"

#Convert columns
gaz <- type_convert(gaz, col_types = cols( 
  FEATURE_ID = col_double(),
  PRIM_LAT_DEC = col_double(),
  PRIM_LONG_DEC = col_double(),
  SOURCE_LAT_DEC = col_double(),
  SOURCE_LONG_DEC = col_double(),
  ELEV_IN_M = col_double(),
  DATE_CREATED = col_date(format = "%m/%d/%Y"),
  DATE_EDITED = col_date(format = "%m/%d/%Y")
))

head(gaz) #Check that conversions worked

```

4. Delete rows with NA for primary longitude or latitude or where the feature is not California.

```{r filter, message = F, warning = F}

gaz <- gaz %>%
  filter(STATE_ALPHA == "CA") %>% 
  filter(PRIM_LAT_DEC != 0, PRIM_LONG_DEC != 0)

```


5. Create a connection to the database and add gaz.

```{r database, message = F, warning = F}

gaz.db <- DBI::dbConnect(RSQLite::SQLite(), path = ":memory:")

copy_to(gaz.db, gaz, "gaz", # What does the "::" mean?
  temporary = FALSE, 
  overwrite = TRUE,
  indexes = list(
    "FEATURE_ID", "FEATURE_NAME" 
  )
)

# Create a table cookie that connects to the database
gaz_db <- tbl(gaz.db, "gaz")
gaz_db

```

## Analyze the Data

1. What is the most frequenty appearing feature name?

```{r feature_name, message = F, warning = F}

# Type convert to a factor
gaz_db$FEATURE_NAME <- as.factor(gaz_db$FEATURE_NAME)

# Count and arrange names in descending order
count_names <- dbGetQuery(gaz.db, "SELECT FEATURE_NAME, COUNT() AS n_names FROM gaz
           GROUP BY FEATURE_NAME
           ORDER BY n_names DESC;")

count_names

```

The feature name that appears the most often is "Church of Christ".

2. What is the least frequently occuring feature class?

```{r feature_class, message = F, warning = F}

count_class <- dbGetQuery(gaz.db, "SELECT FEATURE_CLASS, COUNT() AS n_classes FROM gaz
                          GROUP BY FEATURE_CLASS
                          ORDER BY n_classes;")
count_class

```

The feature classes that appear the least are "Sea" and "Isthmus".

3. What is the approximate center point of each county?

Instead of finding the bounding box and calculating the center point, I chose to take the average latitude and longitude for each county. With such a large sample size for each county, this should be close to center point.

```{r center, warning = F, message = F}

county_center <- dbGetQuery(gaz.db, "SELECT COUNTY_NAME AS 'County Name', AVG(PRIM_LAT_DEC) AS 'Center Latitude', AVG(PRIM_LONG_DEC) AS 'Center Longitude' FROM gaz GROUP BY COUNTY_NAME;")

county_center # Get rid of NA

```

4. What are the fractions of the total number of features in each county that are natural vs. man-made?

```{r feature_type, message = F, warning = F}

# Isolate the unique feature classes
classes <- dbGetQuery(gaz.db,"SELECT FEATURE_CLASS FROM gaz
                      GROUP BY FEATURE_CLASS
                      ORDER BY FEATURE_CLASS;")

classes

# Add classes to the database
copy_to(gaz.db, classes, "classes_db",
  temporary = FALSE, 
  overwrite = TRUE
)

# Create a cookie to talk to the table
classes_db <- tbl(gaz.db, "classes_db")

# Create a new tibble that defines each class as natural or manmade
types <- mutate(classes,
                Class_Type = c("Manmade", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Manmade", "Manmade", "Manmade", "Natural", "Manmade", "Manmade", "Natural", "Manmade", "Manmade", "Natural", "Natural", "Manmade", "Manmade", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Manmade", "Manmade", "Natural", "Natural", "Natural", "Natural", "Natural", "Manmade", "Manmade", "Manmade", "Manmade", "Manmade", "Natural", "Natural", "Manmade", "Manmade", "Natural", "Natural", "Manmade", "Manmade", "Natural", "Manmade", "Natural", "Natural", "Natural", "Natural", "Natural", "Natural", "Manmade", "Manmade", "Manmade", "Natural", "Manmade", "Natural")
)

# Add types to the database
copy_to(gaz.db, types, "types_db",
  temporary = FALSE, 
  overwrite = TRUE
)

# Create a cookie to talk to the table
types_db <- tbl(gaz.db, "types_db")

# Join to gaz, select desired columns
class_types <- dbGetQuery(gaz.db, "SELECT gaz.FEATURE_NAME, gaz.FEATURE_CLASS, gaz.COUNTY_NAME, types_db.Class_Type FROM gaz JOIN types_db ON gaz.FEATURE_CLASS = types_db.FEATURE_CLASS;")

class_types

# Type convert county name to a factor
class_types$COUNTY_NAME <- as.factor(class_types$COUNTY_NAME)

copy_to(gaz.db, class_types, "class_types_db",
  temporary = FALSE, 
  overwrite = TRUE
)

# Create a cookie to talk to the table
class_types_db <- tbl(gaz.db, "class_types_db")
class_types_db

# Pull the different metrics you need
manmade <- dbGetQuery(gaz.db, "SELECT COUNTY_NAME AS 'County', count(Class_Type) AS 'Manmade' FROM class_types_db WHERE Class_Type = 'Manmade' AND COUNTY_NAME != 'NA' GROUP BY COUNTY_NAME;")
manmade

# Add to the database
copy_to(gaz.db, manmade, "manmade",
  temporary = FALSE, 
  overwrite = TRUE
)

# Create a cookie to talk to the table
manmade_db <- tbl(gaz.db, "manmade_db")

natural <- dbGetQuery(gaz.db, "SELECT COUNTY_NAME AS 'County', count(Class_Type) AS 'Natural' FROM class_types_db WHERE Class_Type = 'Natural' AND COUNTY_NAME != 'NA' GROUP BY COUNTY_NAME;")

# Add to the database
copy_to(gaz.db, natural, "natural_db",
  temporary = FALSE, 
  overwrite = TRUE
)

# Create a cookie to talk to the table
natural_db <- tbl(gaz.db, "natural_db")

# Join these and calculate the fractions
fractions <- dbGetQuery(gaz.db, "SELECT manmade_db.County, CAST(manmade_db.Manmade AS REAL)/(manmade_db.Manmade + natural_db.Natural) AS 'Fraction Manmade', CAST(natural_db.Natural AS REAL)/(manmade_db.Manmade + natural_db.Natural) AS 'Fraction Natural' FROM manmade_db JOIN natural_db ON manmade_db.County = natural_db.County;")

fractions

```



